---
title: "plyexperiment"
output: github_document
---


[![Travis build status](https://travis-ci.org/kevinwang09/plyexperiment.svg?branch=master)](https://travis-ci.org/kevinwang09/plyexperiment)

# Introduction 

The S4 object in Bioconductor is great for many things. But every now and then, I have problems manipulating the column or the row data of a S4 object. This package provides several `dplyr`-style functions to manipulate the column data of a S4 object. 

This is an experimental package that only receives occasional updates.


# Demo

Suppose we have a `SingleCellExperiment` object, and we want to add in the cell-type information 
```{r, message = FALSE}
library(SingleCellExperiment)
library(plyexperiment)
library(dplyr)

set.seed(123)
y <- matrix(rnorm(2000), ncol=20)
sce <- SingleCellExperiment(list(counts=y))

cell_type = sample(LETTERS, ncol(y), replace=TRUE)

## Traditional indexing
sce$cell_type = cell_type

## plyexperiment syntax
sce = sce %>% mutate_col(cell_type)
```

The benefit of `plyexperiment` is similar to the benefits of `tidyverse` syntax and benefits from chaining using the pipe operators. For example, the filter function is useful in spliting the data based on information in the `colData`. 

```{r}
sce %>% filter_col(cell_type == "A")
```


## Exporting `reducedDim` data

The `scater` package contains a lot of useful functions in manipulating SingleCellExperiment objects. In particular, the `reducedDim` slot typically stores dimensional reduced data matrices. 

However, to customise the plots based on your own personal taste could be a bit tougher. For example, if you want to overlay the expression of a cell-type marker onto the tSNE plot. 

```{r}
library(scater)
example_sce <- mockSCE()
example_sce <- logNormCounts(example_sce)
example_sce <- runTSNE(example_sce, scale_features=NULL)

plotTSNE(example_sce, colour_by = "Mutation_Status")
```


```{r}
(tsne_tbl = example_sce %>% 
  export_reducedDim(type = "TSNE", 
                    feature = c("Gene_0001", "Gene_0002")))
  
tsne_tbl %>% 
  ggplot(aes(x = V1, y = V2,
             colour = value)) +
  geom_point() +
  facet_wrap(~gene_name)
```


# Acknowledgement

This package was forked over from [`sa-lee/plyexperiment`](https://github.com/sa-lee/plyexperiment) and the main functions were originally written by Davis McCarthy and Aaron Lun. 
